---
title: "Math Project Student Survey Data Cleaning Syntax"
editor_options: 
  chunk_output_type: console
---

## Data cleaning date: 2023-02-15
## Cleaned by: Crystal Lewis
## Project: Math intervention project
## Wave: Wave 1
## Data: Student Survey Data

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

## libraries ##
library(tidyverse)
library(janitor)
library(stringr)
library(labelled)
library(readxl)
library(pointblank)
library(haven)

```

```{r Read in data}

svy <- read_csv("data/w1_mathproj_stu_svy_raw.csv")

```

```{r (01) Review the data}

glimpse(svy)

```

```{r (02) Adjust cases as needed}

# Check for duplicates - 1347 is duplicated

svy %>%
  get_dupes(stu_id)

# Remove duplicates 
# Both surveys are complete - arrange by svy_date ascending

svy <- svy %>%
  arrange(svy_date) %>%
  distinct(stu_id, .keep_all = TRUE)

# Check - Review data after dropping the duplicates

svy

```

```{r (03) De-identify data}

# Drop `svy_date`

svy <- svy %>%
  select(-svy_date)

# Check - Review after drop

names(svy)

```

```{r (07) Normalize variables}

# Fix `math2`

# Check - Review before - there is a return in one value

svy %>%
  tabyl(math2)

# Remove "\n"

svy <- svy %>%
  mutate(math2 = str_remove(math2, "\n"))

# Check - Review after

svy %>%
  tabyl(math2)

```

```{r (09) Update variable types}

# Update both `math2` to numeric

# Check - Review before

class(svy$math2)

# Change variable types - pay close attention to warnings here - if warnings are thrown there may still be characters in the variable

svy <- svy %>%
  mutate(math2 = as.numeric(math2))

# Check - Review after

class(svy$math2)


```

```{r (11) Construct variables}

# Construct `int`

svy <- svy %>%
  mutate(int = case_when(
    stu_id %in% c(1347, 1368, 1387) ~ 1,
    TRUE ~ 0
  ))

# Review `int`

svy %>%
  tabyl(int)


```

```{r (12) Add missing values}

svy <- svy %>% 
  mutate(across(math1:math4, ~replace_na(., -99)))

```

```{r (13) Add metadata}

# Add value labels

svy <- svy %>% 
  mutate(across(math1:math4, 
            ~labelled(., labels = c(`strongly disagree` = 1, 
                                    `disagree` = 2, 
                                    `agree`= 3,
                                    `strongly agree` = 4)))) %>%
  set_value_labels(int = c(treatment = 1, control = 0))

# Check - value labels

val_labels(svy)

# Add missing value labels

svy <- svy %>% 
  mutate(across(math1:math4, ~(`na_values<-`
                                (., -99))))

# Check - missing value labels

na_values(svy)

# Add variable labels using our data dictionary

# Read in data dictionary and pull variable labels into a named list

var_labels <- read_excel("mathproj_stu_svy_data-dictionary.xlsx") %>%
  select(var_name, label) %>%
  deframe() %>%
  as.list()

# Add variable labels

svy <- svy %>%
  set_variable_labels(.labels = var_labels)

# Check- variable labels

var_label(svy)


```

```{r (14) Validate data}

# Validate data using the `pointblank` package

create_agent(svy) %>%
  rows_distinct(columns = vars(stu_id)) %>%
  col_vals_between(columns = c(stu_id), left = 1300, right = 1400, na_pass = TRUE) %>%
  col_vals_in_set(columns = c(grade_level), 
                  set = c(9, 10, 11, 12, NA)) %>%
  col_vals_in_set(columns = c(int),
                  set = c(0, 1, NA)) %>%
  col_vals_in_set(columns = c(math1:math4),
                  set = c(1, 2, 3, 4, NA)) %>%
  interrogate()

```

```{r Export data}

write_sav(svy, "data/w1_mathproj_stu_svy_clean.sav")

```

```{r Create session info}

sessionInfo()

```

R version 4.1.2 (2021-11-01)
Platform: x86_64-w64-mingw32/x64 (64-bit)
Running under: Windows 10 x64 (build 19045)

Matrix products: default

locale:
[1] LC_COLLATE=English_United States.1252  LC_CTYPE=English_United States.1252   
[3] LC_MONETARY=English_United States.1252 LC_NUMERIC=C                          
[5] LC_TIME=English_United States.1252    

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] haven_2.5.0            pointblank_0.10.0.9000 readxl_1.4.0           labelled_2.9.1        
 [5] janitor_2.1.0          forcats_0.5.1          stringr_1.5.0          dplyr_1.0.9           
 [9] purrr_0.3.4            readr_2.1.2            tidyr_1.1.4            tibble_3.1.8          
[13] ggplot2_3.4.0          tidyverse_1.3.2       

loaded via a namespace (and not attached):
 [1] lubridate_1.8.0     assertthat_0.2.1    digest_0.6.29       utf8_1.2.2          R6_2.5.1           
 [6] cellranger_1.1.0    backports_1.4.1     reprex_2.0.1        evaluate_0.20       httr_1.4.3         
[11] pillar_1.8.1        rlang_1.0.6         googlesheets4_1.0.1 rstudioapi_0.13     checkmate_2.1.0    
[16] rmarkdown_2.16      googledrive_2.0.0   bit_4.0.4           munsell_0.5.0       broom_1.0.3        
[21] compiler_4.1.2      modelr_0.1.10       xfun_0.30           pkgconfig_2.0.3     htmltools_0.5.3    
[26] tidyselect_1.2.0    fansi_1.0.3         crayon_1.5.2        tzdb_0.3.0          dbplyr_2.1.1       
[31] withr_2.5.0         commonmark_1.8.0    grid_4.1.2          jsonlite_1.8.4      gtable_0.3.1       
[36] lifecycle_1.0.3     DBI_1.1.2           magrittr_2.0.3      scales_1.2.1        cli_3.2.0          
[41] stringi_1.7.12      vroom_1.5.7         fs_1.5.2            snakecase_0.11.0    xml2_1.3.3         
[46] ellipsis_0.3.2      generics_0.1.3      vctrs_0.5.2         tools_4.1.2         bit64_4.0.5        
[51] glue_1.6.2          hms_1.1.2           yaml_2.3.5          parallel_4.1.2      fastmap_1.1.0      
[56] colorspace_2.0-2    gargle_1.2.0        gt_0.5.0            rvest_1.0.2         blastula_0.3.2     
[61] knitr_1.40          sass_0.4.2         